Community biomass and species richness temporal trends reveal contrasting assembly and disassembly processes in natural food-webs

### 2024-08-07 paper/manuscript.Rmd
Revise discussion + add theoretical model methods

rm(full_data2)
```

@@

```{r, results=FALSE}
tar_load(glob_tps_trends_decade_no_drivers)
gbm <- get_effect_ci(resp = "log_bm_std", ci_lvl = "level:0.80", p = TRUE)
gr <- get_effect_ci(resp = "log_rich_std", ci_lvl = "level:0.80", p = TRUE)
gt <- get_effect_ci(resp = "w_trph_lvl_avg", ci_lvl = "level:0.80", p = FALSE)
gct <- get_effect_ci(resp = "connectance", ci_lvl = "level:0.80", p = FALSE)
gl <- get_effect_ci(resp = "prop_redundant_links", ci_lvl = "level:0.80", p = FALSE)
```

@@

We further tested if we find the same patterns using spatial variations in
food-web structure rather than temporal changes over time, i.e. if the
space-for-time hypothesis hold. To do so, we used the same dataset and same
structural equation model but with the median food-web structure per site
instead of the temporal trends. We additionally performed theoretical
simulations with a bioenergetic food-web model to further get theoretical
expectations about how food-web structure changes according to community biomass and
species richness variations (see Methods). As in temporal data, species richness
had a positive effect on community biomass both in empirical data and in
theoretical simulations (resp. $r_\delta =$ `r rdsp$bm$r` and `r rdm$bm$r`, Fig.
\@ref(fig:std) and S5). In both empirical data and theoretical simulations, community
biomass had a positive effect on connectance (resp. $r_\delta =$ `r rdsp$ct$bm`
and `r rdm$ct$bm`, Fig. \@ref(fig:std)), average trophic level (resp.
$r_\delta =$ `r rdsp$tlvl$bm` and `r rdm$tlvl$bm`), and redundancy (resp.
$r_\delta =$ `r rdsp$red$bm` and `r rdm$red$bm`), as in temporal analysis.

The effects of species richness were more complex, because species richness had
indirect effects on food-web structure through community biomass. Temporal
trends in species richness had a direct negative effect on connectance
($r_\delta =$ `r rdtps$ct$r`, Fig. \@ref(fig:sem)f), a positive
indirect effect through community biomass ($r_\delta =$ `r ridtps$ct$r`), so
that species richness temporal trends had a negative effect in total on temporal in
connectance ($r_\delta =$ `r rttps$ct$r`, Fig. \@ref(fig:std)).
In contrast, temporal trends in species had no direct effects on the temporal
trends in average trophic level ($r_\delta =$ `r rdtps$tlvl$r`, Fig.
\@ref(fig:sem)g), a positive indirect effect through community
biomass ($r_\delta =$ `r ridtps$tlvl$r`), which adds up to a positive total
effect of species richness temporal trends on average trophic level ($r_\delta
=$ `r rttps$tlvl$r`, Fig. \@ref(fig:std)). Species richness temporal trends had
direct and indirect positive effect on the temporal trends in redundancy (resp.
$r_\delta =$ `r rdtps$red$r`, `r ridtps$red$r`), which adds up to a strong
positive effect $r_\delta =$ `r rttps$red$r`. Furthermore, we found that
temporal trends of the proportion of piscivorous species was negatively
associated with the temporal trends of species richness (Fig. S3), meaning that
communities that increase in richness over time have an increase in the
proportion of non-piscivorous species over time and vice versa. It means
food-webs that have experienced a decrease in species richness over time have
decreased in average trophic level and redundancy but increase in connectance,
and conversely, those that increased in species richness have increased in
average trophic level, redundancy, but decreased in connectance.

The effects of spatial variations in species richness on food-web structure
mostly agreed with the ones found with temporal trends. We found a direct negative
effect of species richness on connectance both using data and model simulations
(resp. $r_\delta =$ `r rdsp$ct$r` and `r rdm$ct$r`, Fig. \@ref(fig:std)), on
average trophic level (resp. $r_\delta =$ `r rdsp$tlvl$r` and `r rdm$tlvl$r`),
but a direct positive effect on redundancy (resp. $r_\delta =$ `r rdsp$red$r`
and `r rdm$red$r`). Species richess had indirect positive effect of species
richness through community biomass, which added up to strong total negative
effects on connectance in empirical data and simulations (resp. $r_\delta =$
`r rtsp$ct$r` and `r rtm$ct$r` ) and positive on redundancy (resp. $r_\delta =$
`r rtsp$red$r` and `r rtm$red$r`). We found that species richness had a total
positive effect on average trophic level in bioenergetic model, in agreement
with temporal trends (resp. $r_\delta =$ `r rtm$tlvl$r`), but a total negative
effect in spatial variations (resp. $r_\delta =$ `r rtsp$tlvl$r`).

@@

We found remarkable heterogeneity in community biomass and species richness
changes over the last decades in riverine fish communities, linked to changes in
food-web structure. Community biomass changes spanned from `r grbmn`% to `r
grbmx`% per decade, species richness from `r grrhn`% to +`r grrhx`% per decade,
in line previous reports of the heterogeneity of local biodiversity changes in
terrestrial, marine and freshwater realms[@dornelas_assemblage_2014;
@blowes_geography_2019; @danet_past_2024]. In turn, connectance temporal changes
spanned from `r grctn` to +`r grctx` per decade and average trophic level, from
`r grtln` to +`r grtlx` per decade. We questioned the effects of community
biomass and species richness changes over time on temporal changes in food-web
structure, that we also tested across space and with simulations using a
bioenergetic food-web model. We found positive effects of species richness
changes on changes in community biomass that were consistent across temporal and
spatial analysis and with the theoretical model. The effects of community
biomass changes on food-web structure were also consistent across analysis,
suggesting that loss of community biomass are linked to the loss of the trophic
species of high trophic level. However, changes in species richness had
contrasting effects on food-web structure depending on the type of analysis,
temporal or spatial and theoretical simulations. It suggests that food-web
structural changes induced by changes in species richness differs from the one
induced by the changes in community biomass, and that they imply different
processes across time and space. However, all analysis converged to indicate
that community biomass and species richness declines trigger a decrease in
food-web redundancy. In what follows, we discuss how changes in community
biomass and species richness changes over time and space are related to food-web
assembly and disassembly processes.

@@

Species richness temporal changes were positively associated to changes in average
trophic level, such as food-webs that have suffered a decline in species richness
have experienced a concomitant decline in average trophic level, those have seen
an increase in species richness have experienced an increase in average trophic
level. However, the effect of species richness temporal trends on average
trophic level were twice lower than the one of community biomass, and had no
effects on connectance temporal changes. Altogether, it suggests that species
richness increase and decrease were less clearly linked to specific changes in
food-web structure. Species richness temporal trends could be more linked to
changes in the redundancy of trophic roles in the food-webs, such as it has
little effects of food-web structure but increases its community biomass. This
interpretation align with the insurance hypothesis [@yachi_biodiversity_1999;
@loreau_biodiversity_2021], such as species richness and associated redundancy
in species roles can increase ecosystem stability by buffering the impacts of
perturbations [@sanders_trophic_2018].

@@

## Conclusion and perspectives

- Revise: biomass pyramids, connectance different from top predators
- Add redundancy

We investigated how food-web assemble and disassemble according to changes in
community biomass and species richness. Our study shows that temporal changes in
community biomass induced changes in food-web structure, such as decline in
biomass translates in food-webs with *lower connectance* and fewer top predators
while increases in biomass results in food-webs with higher connectance and more
top predators. In turn, decline in species richness translates in food-webs with
fewer trophic links and top predators while
increases in biomass results in food-webs with more trophic links and top
predatorstranslate in low changes in
food-web structure, suggesting that species richness changes relates more to
trophic role
redundancy than in changes in food-web structure. Using replicating the analysis
over spatial gradient and with a theoretical model, our study shows that
community biomass changes over space and theoretical model results in the same
changes in food-web structure than in temporal analysis, which is not the case
for species richness. It suggests that using space-for-time substitution to
infer assembly and disassembly processes in food-web structure works well for
community biomass but not so well for species richness.

In the context of biodiversity crisis, our study reveals that declines of
community biomass results in food-web downgrading. As species of top trophic
levels plays a great role in ecosystem functioning [@estes_trophic_2011;
@wang_biodiversity_2018; @thebault_food-web_2003], it raises concerns about the
functioning of such degraded ecosystems. Loss of species richness in turn did
not affect significantly food-web structure, suggesting that it implies loss of
redundancy in trophic roles. As redundancy has a key role in buffering ecosystem
variability against perturbations [@sanders_trophic_2018;
@yachi_biodiversity_1999], it raises concerns about the stability of such
food-webs. We used here empirical freshwater food-web inferred from in-situ fish
bodysize and diet [@danet_species_2021; @bonnaffe_comparison_2021], which do not
account for the variability in invertebrates and primary producers presences,
and do not a number of processes such as changes in interaction strengths.
However, the theoretical model embedding complete food-webs showed remarkably
similar results to the empirical data.

@@

At each site, yearly food-webs were described by their community biomass and
species richness. Yearly food-web structure was described by their connectance,
describing the degree of generalism in a network [@dunne_network_2006;
@delmas_analysing_2019]. It was computed as the ratio between the number of
actual links and the number of potential links ($C = \frac{L}{S(S-1)}$, with $L$
being the number of links and $S$ the number of trophic species). We also
computed the average trophic level, weighted by the biomass of the trophic
species. The trophic level of a trophic species was computed recursively as 1
added to the mean trophic level of the preys that it consumes
[@herendeen_ecological_2008].

@@

We modeled the relationships among food-web structure, that is
connectance and average trophic level, and community biomass ($dB$) and species
richness ($dR$) with Structural Equation Models. We modeled species
richness as related to community biomass and food-web structure. In
turn community biomass was related to food-web structure. We computed one
structural equation model to assess the relationships among the temporal trends
and one for the relationships over the spatial gradient. As spatial variation of
food-web structure are partly driven by environmental gradients, so we
controlled relationships among variables by adding two PCA axis
synthesizing respectively (1) stream gradient from upstream to downstream and
(2) enrichment [details in @danet_species_2021]. Enrichment was characterized by
the average Biological Oxygen Demand [BOD, @mallin_factors_2006]. We further
added a random effect of the hydrographic basin on the intercept.

@@

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were
standardized and the confidence intervals were computed with semEff R package
[@murphy_semeff_2021]. Confidence intervals were obtained by bootstrapping the
linear models 100 times and computing adjusted bootstrap percentile interval
(BCa), which are adjusted for the asymmetric distribution of resampling. The marginal
and conditional $R^2$ associated to the linear models of the SEMs were computed
with the Nagelkerke method [@lefcheck_piecewisesem_2016, Fig. S5].

@@

>>>>

rm(full_data2)
tar_load(glob_tps_trends_decade_no_drivers)
```

@@

```{r, results=FALSE}
gbm <- get_effect_ci(resp = "log_bm_std", ci_lvl = "level:0.80", p = TRUE)
gr <- get_effect_ci(resp = "log_rich_std", ci_lvl = "level:0.80", p = TRUE)
gt <- get_effect_ci(resp = "w_trph_lvl_avg", ci_lvl = "level:0.80", p = FALSE)
gct <- get_effect_ci(resp = "connectance", ci_lvl = "level:0.80", p = FALSE)
gl <- get_effect_ci(resp = "prop_redundant_links", ci_lvl = "level:0.80", p = FALSE)
```

@@

We further tested if we find the same patterns using spatial variations in
food-web structure rather than temporal changes over time, i.e. if the
space-for-time hypothesis hold. To do so, we used the same dataset and same
structural equation model but with the median food-web structure per site
instead of the temporal trends. We additionally performed theoretical
simulations with a bioenergetic food-web model to further get theoretical
expectations about how food-web structure changes according to community biomass and
species richness variations (see Methods). As in temporal data, species richness
had a positive effect on community biomass both in empirical data and in
theoretical simulations (resp. $r_\delta =$ `r rdsp$bm$r` and `r rdm$bm$r`, Fig.
\@ref(fig:std) and S5). In both empirical data and theoretical simulations, community
biomass had a positive effect average trophic level (resp.
$r_\delta =$ `r rdsp$tlvl$bm` and `r rdm$tlvl$bm`), and redundancy (resp.
$r_\delta =$ `r rdsp$red$bm` and `r rdm$red$bm`), as in temporal analysis.
We found weak evidence of a positive effect of spatial variation in community
biomass on connectance in empirical data ($r_\delta =$ `r rdsp$ct$bm`) but
strong evidence in theoretical simulation ($r_\delta =$ `r rdm$ct$bm`, Fig. \@ref(fig:std)).

The effects of species richness were more complex, because species richness had
indirect effects on food-web structure through community biomass. Temporal
trends in species richness had a direct negative effect on connectance
($r_\delta =$ `r rdtps$ct$r`, Fig. \@ref(fig:sem)f), a positive
indirect effect through community biomass ($r_\delta =$ `r ridtps$ct$r`), so
that species richness temporal trends had a negative effect in total on temporal in
connectance ($r_\delta =$ `r rttps$ct$r`, Fig. \@ref(fig:std)).
In contrast, temporal trends in species had no direct effects on the temporal
trends in average trophic level ($r_\delta =$ `r rdtps$tlvl$r`, Fig.
\@ref(fig:sem)g), a positive indirect effect through community
biomass ($r_\delta =$ `r ridtps$tlvl$r`), which adds up to a positive total
effect of species richness temporal trends on average trophic level ($r_\delta
=$ `r rttps$tlvl$r`, Fig. \@ref(fig:std)). Species richness temporal trends had
direct and indirect positive effect on the temporal trends in redundancy (resp.
$r_\delta =$ `r rdtps$red$r`, `r ridtps$red$r`), which added up to a strong
positive effect ($r_\delta =$ `r rttps$red$r`). Furthermore, we found that
temporal trends of the proportion of piscivorous species was negatively
associated with the temporal trends of species richness (Fig. S3), meaning that
communities that increase in richness over time have an increase in the
proportion of non-piscivorous species over time and vice versa. It means
food-webs that have experienced declines in species richness over the last
decades have decreased in average trophic level and redundancy but increase in
connectance, and conversely, those that experienced increases in species
richness have increased in average trophic level, redundancy, but decreased in
connectance.

The effects of spatial variations in species richness on food-web structure
mostly agreed with the ones found with temporal trends, expect for average
trophic level. We found a direct negative effect of species richness on
connectance both using data and model simulations
(resp. $r_\delta =$ `r rdsp$ct$r` and `r rdm$ct$r`, Fig. \@ref(fig:std)), on
average trophic level (resp. $r_\delta =$ `r rdsp$tlvl$r` and `r rdm$tlvl$r`),
but a direct positive effect on redundancy (resp. $r_\delta =$ `r rdsp$red$r`
and `r rdm$red$r`). Species richness had indirect positive effect of species
richness through community biomass, which added up to strong total negative
effects on connectance in empirical data and simulations (resp. $r_\delta =$
`r rtsp$ct$r` and `r rtm$ct$r` ) and positive on redundancy (resp. $r_\delta =$
`r rtsp$red$r` and `r rtm$red$r`). We found that species richness had a total
positive effect on average trophic level in bioenergetic model, in agreement
with temporal trends (resp. $r_\delta =$ `r rtm$tlvl$r`), but a total negative
effect in spatial variations (resp. $r_\delta =$ `r rtsp$tlvl$r`).

@@

We found remarkable heterogeneity in community biomass and species richness
changes over the last decades in riverine fish communities, linked to changes in
food-web structure. Community biomass changes spanned from -`r grbmn`% to `r
grbmx`% per decade, species richness from `r grrhn`% to +`r grrhx`% per decade,
in line previous reports of the heterogeneity of local biodiversity changes in
terrestrial, marine and freshwater realms[@dornelas_assemblage_2014;
@blowes_geography_2019; @danet_past_2024]. In turn, connectance temporal changes
spanned from `r grctn` to +`r grctx` per decade, average trophic level, from
`r grtln` to +`r grtlx` per decade and the proportion of redundant trophic links
from `r grln` to +`r grlx` per decades. We questioned the effects of community
biomass and species richness changes over time on temporal changes in food-web
structure, that we also tested across space and with simulations using a
bioenergetic food-web model. We found positive effects of species richness
changes on changes in community biomass that were consistent across temporal and
spatial analysis and with the theoretical model. The effects of community
biomass changes on food-web structure were also consistent across analysis,
suggesting that loss of community biomass are linked to the loss of the trophic
species of high trophic level. However, changes in species richness had
contrasting effects on food-web structure depending on the type of analysis,
temporal or spatial and theoretical simulations. It suggests that food-web
structural changes induced by changes in species richness differs from the one
induced by the changes in community biomass, and that they imply different
processes across time and space. However, all analysis converged to indicate
that community biomass and species richness declines trigger a decrease in
food-web redundancy. In what follows, we discuss how changes in community
biomass and species richness changes over time and space are related to food-web
assembly and disassembly processes.

@@

Species richness temporal changes were positively associated to changes in
average trophic level through community biomass and in connectance mainly
through direct effects, such as food-webs that have suffered a decline in
species richness over last decades have experienced a concomitant decline in
average trophic level and increase in connectance, those have seen an increase
in species richness have experienced an increase in average trophic level and
decline in connectance. However, the effect of species richness temporal trends
on average trophic level were twice lower than the one of community biomass.
Altogether, it suggests that species richness increase and decrease were less
clearly linked to specific changes in top trophic levels than community biomass.
Species richness temporal trends could be more linked to changes in the
redundancy of trophic roles in the food-webs, such as it has little effects of
food-web structure but increases its community biomass. This interpretation
align with the insurance hypothesis [@yachi_biodiversity_1999;
@loreau_biodiversity_2021], such as species richness and associated redundancy
in species roles can increase ecosystem stability by buffering the impacts of
perturbations [@sanders_trophic_2018].

@@

Our study demonstrates that redundancy of trophic pathways in food-webs are
positively associated to both community biomass and species richness changes, so
as declines in community biomass and species richness triggers a loss in the
proportion of redundant links in food-webs, while increases in community biomass
and species richness over time results in the increase in the proportion of
redundant links in food-webs. Redundant links are the trophic links that can be
removed with triggering secondary extinctions [@allesina_who_2004;
@bodini_using_2009], such an higher proportion of redundant interactions
increase robustness of food-webs to perturbations [@allesina_functional_2009].
It suggests that declines in community biomass and species richness result in a
decline of food-web robustness to further perturbations.

Our study brings evidences that temporal changes in community biomass and
species richness are related to changes in food-web structure in empirical
settings. We completed the temporal analysis with a spatial analysis and
theoretical simulations. Altogether, it suggests that community biomass changes
are associated with changes in the presence of top predators, in agreement with
the hypothesis that food-web maximum trophic level are limited by the amount of
energy available in lower trophic level. In contrast, species richness changes
seems to be linked to changes in upper and lower part of the food-webs.
Importantly, temporal and spatial analysis using empirical fish food-webs are
largely in agreement with theoretical simulations, which highlight the potential
generality of the food-web assembly rules that we report. The study finally
highlights that increases in community biomass and species brings robustness to
food-webs by increasing the proportion of redundant trophic links. Conversely, it
suggests that community declines in species richness and community biomass are
likely to decrease significantly the community robustness to further
perturbations. Thereby, the numerous reported declines in biodiversity might
hide increase in the fragility of ecosystems.

@@

At each site, yearly food-webs were described by their community biomass and
species richness. Yearly food-web structure was described by their connectance,
describing the degree of generalism in a network [@dunne_network_2006;
@delmas_analysing_2019]. It was computed as the ratio between the number of
actual links and the number of potential links ($C = \frac{L}{S(S-1)}$, with $L$
being the number of links and $S$ the number of trophic species). We also
computed the average trophic level, weighted by the biomass of the trophic
species. The trophic level of a trophic species was computed recursively as 1
added to the mean trophic level of the preys that it consumes
[@herendeen_ecological_2008]. We computed the number of redundant and functional
links using the dominator tree approach [@allesina_functional_2009]. A species
$i$ is a dominator of species $j$ if $i$ mediates the connection $j$ to basal
resources such as the extinction of $i$ remove the connection of $j$ to basal
resources. It is then to identify the set of trophic links that are essential to
connect species to basal resources, i.e. functional links, and the
redundant links. This method requires a single basal root in the food-webs. To
do so, we added a basal dummy resource node to the empirical food-webs connected
to biofilm and phytoplankton resource node. For food-webs generated from
theoretical simulations (see below), we added a resource node connected to all
primary producers. We computed the set of dominator trees using the
`dominator_tree()` function in `igraph` R package. We computed the number of
functional links as the number of links in the resulting dominator tree of the
food-web. We checked that the procedure reproduced the examples presented in
@allesina_functional_2009. We then derived the proportion of redundant links in
food-webs of empirical and theoretical simulations.

@@

We modeled the relationships among food-web structure, that is
connectance, average trophic level, the proportion of redundant trophic links, and
community biomass ($dB$) and species richness ($dR$) with Structural Equation
Models. We modeled species
richness as related to community biomass and food-web structure. In
turn community biomass was related to food-web structure. We computed one
structural equation model to assess the relationships among the temporal trends
and one for the relationships over the spatial gradient, we added a random effect
of the hydrographic basin on the intercept for the latter one.

@@

The coefficients of the Structural Equation Models were computed with
PiecewiseSEM R package [@lefcheck_piecewisesem_2016]. The coefficients were
standardized and the confidence intervals were computed with semEff R package
[@murphy_semeff_2021]. Confidence intervals were obtained by bootstrapping the
linear models 1000 times and computing adjusted bootstrap percentile interval
(BCa), which are adjusted for the asymmetric distribution of resampling. The marginal
and conditional $R^2$ associated to the linear models of the SEMs were computed
with the Nagelkerke method [@lefcheck_piecewisesem_2016, Fig. S5].

@@

## Theoretical simulations

```{r}
tar_load(sim_param)
nsim <- nrow(sim_param)
nfw <- length(unique(sim_param$fw_id))
s_range <- c(min = min(sim_param$S), max = max(sim_param$S))
c_range <- c(min = min(sim_param$C), max = max(sim_param$C))
k_range <- c(min = min(sim_param$K), max = max(sim_param$K))
```

We used a bioenergetic food-web model to simulate the effects of community
biomass and species richness changes on food-web structure. This food-web model
lifts one of the limitation of our data, that is it includes the biomass of all
the food-webs from primary producers to top predators. This model allows to
simulate complex communities because its parametrization depends mainly on
species bodymasses. We first generated `r nfw` food-webs varying in species
richness (range: `r s_range["min"]`-`r s_range["max"]`)
and connectance (range: `r c_range["min"]`-`r c_range["max"]`) using the niche
model algorithm, which reproduces food-web topology tipically found in empirical
food-webs [@williams_simple_2000]. We set species bodymasses ($M_i$) using a predator-prey
mass ratio of 100 ($Z$) [@brose_allometric_2006; @brose_predator_2019], such as $M_i =
Z^{(TL_i - 1)}$, $TL_i$ being the trophic level of the species $i$. The
metabolic rates of the species ($x_i$) were then derived from allometric
relationships between metabolism and bodymass empirically estimated
[@brose_allometric_2006], such as $x_i = aM_i^b$, with $a = xx$ and $b = -0.25$.

The dynamics of species biomasses are described by an ODE systems of two
equations, for primary producers (eq. \@ref(eq:prod)) and consumers (eq. \@ref(eq:cons)) respectively:

\begin{equation}
    (\#eq:prod)
	\frac{dB_i}{dt} =
    r_i B_i G_i
    - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
\end{equation}

\begin{equation}
    (\#eq:growth)
    G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
\end{equation}

The primary producer biomass increase with a growth rate ($r_i$) of 1 and a
logistic growth (eq. \@ref(eq:growth)), growth being limited by intraspecific
($\alpha_{ii}$, set to 1), interspecific ($\alpha_{ij}$, set to 0.5) competition,
and carrying capacity ($K_i$, varying in the simulations). Primary producers
loses biomass by being eaten by consumers at a rate that depend on consumer
metabolic rates ($x_j$), attack rates ($y_j$, set to 8 @delmas_analysing_2019),
a functional response (F_{ji}), and assimilation efficiency ($e_{ij}$, set to
0.45 for herbivores and 0.85 for carnivores, @delmas_analysing_2019).

\begin{equation}
    (\#eq:cons)
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i
\end{equation}

Consumers gain biomass by eating other species and lose biomass by being
consumed and by metabolic loses ($x_i$, eq. \@ref(eq:cons)). The functional
response (eq. \@ref(eq:holling)) of consumer feeding depends on the preference
for its preys ($\omega_{ij}$, set even across consumer preys) and
half-saturation rate ($B_0$, set to 0.5). The hill coefficient ($h$, set to 1.5)
determines the shape of the functional response (Holling's type II or III).

\begin{equation}
    (\#eq:holling)
    F_{ij} = \frac{\omega_{ij} B_j^h}
    {B_0^h + \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
\end{equation}

We simulated the food-webs dynamics across a gradient of global carrying capacity (K,
range: `r k_range["min"]`-`r k_range["max"]`). The carrying capacity were
standardized according to the number of producers and interspecific competition
($K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K$) to prevent changes in productivity
due to system size [@ives_stability_2000; @loreau_synchrony_2008]. We ran
simulations for 5000 timesteps. At the end of simulations, we removed dead
species and measured community biomass, species richness and food-web structure.
If the final community contained disconnected producers (i.e. without consumers)
or disconnected consumers (i.e. consumers without preys), we discarded them and
run the simulations again with the same parametrisation. The simulations were
done using the Julia package `EcologicalNetworksDynamics.jl`
[@lajaiiti_ecological_2024].
